---
title: "Assignment 8: R notebooks and gapminder w/ggplot"
output: html_notebook
---
## Loading libraries
```{r}
library(gapminder)
library(tidyverse)
```

## Creating a gapminder data frame
```{r}
data("gapminder")
gapminder

summary(gapminder)
```

## Isolating a column "gdpPercap"
```{r}
x <- mean(gapminder$gdpPercap)
x

attach(gapminder)
```

## Getting median and plotting histogram of life expectancy
```{r}
median(pop)

hist(lifeExp)
hist(pop) 
hist(log(pop))
```

## Boxplots
```{r}
boxplot(lifeExp ~ continent)
plot(lifeExp ~ gdpPercap)
plot(lifeExp ~ log(gdpPercap))
```

## New naming for the dataframe and getting median
```{r}
mygap <- gapminder
summary(mygap)
x_mygap <- mean(mygap$gdpPercap)
median(mygap$pop)
```

## Histograms
```{r}
hist(mygap$lifeExp)
hist(mygap$pop) 
hist(log(mygap$pop))
```

## Boxplots
```{r}
boxplot(lifeExp ~ continent, data=mygap)
plot(lifeExp ~ gdpPercap, data=mygap)
plot(lifeExp ~ log(gdpPercap), data=mygap)
```

## dplyr manipulations and ggplots
```{r}
mygap %>%
  select(country, lifeExp) %>% 
  filter(country == "South Africa" |
           country == "Ireland") %>% 
  group_by(country) %>% 
  summarise(Average_life = mean(lifeExp))
```


```{r}
df1 <- gapminder %>%
  select(country, lifeExp) %>% 
  filter(country == "South Africa" |
           country == "Ireland")
```

## t-test to compare the life expectancy between South Africa and Ireland
```{r}
t.test(data = df1, lifeExp ~ country)
```

## Basic scatterplot aka "geom_point" 
```{r}
mygap %>% 
  filter(gdpPercap < 50000) %>% 
  ggplot(aes(x=gdpPercap, y=lifeExp)) +
  geom_point()
```

## Color-enhanced single plot
```{r}
mygap %>% 
  filter(gdpPercap < 50000) %>% 
  ggplot(aes(x=log(gdpPercap), y=lifeExp, col=continent, size=pop)) +
  geom_point(alpha=0.3) +
  geom_smooth(method=lm)
```

## Continent-faceted scatterplots
```{r}
mygap %>% 
  filter(gdpPercap < 50000) %>% 
  ggplot(aes(x=log(gdpPercap), y=lifeExp, col=year, size=pop)) +
  geom_point(alpha=0.3) +
  geom_smooth(method=lm) +
  facet_wrap(~continent)
```

## Linear regression models
```{r}
lm(lifeExp ~ gdpPercap, data=mygap)
```


```{r}
lm(lifeExp ~ gdpPercap, data=mygap) %>% 
  summary
```


```{r}
lm(lifeExp ~ gdpPercap+pop, data=mygap) %>% 
  summary
```


```{r}
lm(lifeExp ~ log(gdpPercap)+log(pop), data=mygap) %>% 
  summary
```

